DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-04 rebuttal.tex

\begin{mdframed}[style=comment]
This kind of maps can be created using the gravitational lensing technique [see 29, for review]

- It should be "These" and not "This". "These kind of maps can be created using the gravitational lensing technique [see 29, for review]"
\end{mdframed}
Thank you for spotting this mistake. We updated the manuscript.

\begin{mdframed}[style=comment]
A CNN is composed of the following main building blocks [31]: (i) a convolution, (ii) a non-linearity, (iii) a down-sampling operation, (iv) a pooling operation, and (v), optionally, normalization.

- I think that this is generalising a CNN a bit too much. In my mind (and in my personal uses of CNNs) steps (iii) and (iv) are not always necessary either. Perhaps you want to say that you wish to tackle common tasks which often arise in CNNs: i) convolution, (ii) non-linearity, (iii) down-sampling operation, (iv) pooling operation, and (v), optionally, normalisation.
\end{mdframed}
\todo{assigned: @michael}
\nati{I do not know what to say. I just do not care what is the definition of a CNN.}

@@

\begin{mdframed}[style=comment]
Likewise, down-sampling can be achieved by taking one pixel out of n.

- I'm afraid I don't understand what this means at all, could you rephrase it?
\end{mdframed}
This has been rephrased as:
\begin{mdframed}[style=manuscript]
Likewise, down-sampling can be achieved by keeping one pixel every $n$ pixels with $n=2$ being very common.
\end{mdframed}

\begin{mdframed}[style=comment]
A rhombus is a quadrilateral whose four sides all have the same length.

- I'm not sure that this really needs to be said.
\end{mdframed}
We agree with the reviewer and updated the manuscript accordingly.
% \todo{assigned: @michael}

@@

\begin{mdframed}[style=comment]
Similarly as (2), the convolution of the signal $f$ by a kernel $g$ is the scalar product of $f$ with translated versions $T_ig$ of the kernel $g$.

- I think this should be "As with equation (2)" or "Similarly, as in equation (2)" but I can't tell which. Also, it would be really useful if you referred to equations by "equation (x)" rather than just "(x)" throughout - much easier to read that way.
\end{mdframed}
% \todo{Assign: @michael, @tomek}
% \nati{I do not know what we should do.}
Thank you for your suggestion. Ideally we would replace equation references such as "(1)" with "Equation (1)", but given the large number of equation references, we would prefer to stick with a short-form to avoid lengthening the paper further.

\begin{mdframed}[style=comment]
The 1-neighborhood of a node is the set of nodes that are directly connected to it. The k-neighborhood of a node is the set of nodes that are connected to it through paths of length k.

- I'm a little confused here, I think the neighbourhood is every pixel which is closer than the distance of some pixel. The value of k would then be the length of the set of ordered distances for all pixels up to this pixel. Is that correct? For example, take a 3x3 square (on Euclidean space for simplicity), if the central pixel has coordinates (0, 0) then the 1-neighbourhood would be all pixels at a length of 1 pixel or less, i.e. (-1, 0), (+1, 0), (0, -1) and (0, +1) who are all at the same distance so the set of distances is {1} pixel and the length of the set is 1. The 2-neighbourhood would then be all pixels which were a distance $\sqrt{2}$ pixels or less away, i.e. (-1, 0), (+1, 0), (0, -1), (0, +1), (-1, -1), (-1, +1), (+1, -1), (+1, +1) so the set of distances would be {1, $\sqrt{2}$} pixels and the length of the set is 2. Is this correct? If so k is not a length. If this is not correct, I think I misunderstand something and so this should be updated in the text.
\end{mdframed}
\todo{To check: @michael}

Here, the notion of neighboorhood is defined in the context of graphs. As there is no coordinates to compute distances in this context, one way to define the distance between two vertices is to count the number of edges in the shortest path between these two vertices. To clarify this point, we added more context in the manuscript:

\begin{mdframed}[style=manuscript]
In the context of graph, one way to define the distance between two vertices is to count the number of edges in the shortest path between these two vertices. Hence, we define the 1-neighborhood of a vertex as the set of vertices that are directly connected to it. The $k$-neighborhood of a vertex is the set of vertices that are connected to it through paths of length $k$.
\end{mdframed}

@@

\begin{mdframed}[style=comment]
The rectified linear unit (ReLU) $\sigma(\cdot) = \textrm{max}(\cdot, 0)$ is a common choice.

- Is this the choice you adopt? You should mention that you do, if you indeed do.
\end{mdframed}
We followed the reviewer suggestion.

\begin{mdframed}[style=comment]
Note that the output $\bf{Y}\in\mathbb{R}^{N_{pix}\times F_{out}}$ of the last GC (or the output $\bf{Y}\in\mathbb{R}^{N_{stat}\times F_{out}}$ of the ST), ...

- What is ST, you don't mention it before and only once after?
\end{mdframed}
We thank the reviewer for spotting this. Indeed $ST$ was not defined. In the manuscript, we have replaced $ST$ by statistical layer.

@@

\begin{mdframed}[style=comment]
We derived this rule from the Chebyshev polynomial equations and some empirical experiments.

- This is a bit loose, can you explain further and add a citation if you didn't do the experiments yourself?
\end{mdframed}
% \todo{Assign: @nati, @michael}
% \nati{We would need an appendix to do that. It is the last figure of the figure’s notebook (expected filter). I could spend two hours and write some equation... @michael, what do you think?}
We updated the section with a citation and some extra information.
\begin{mdframed}[style=manuscript]
We derived this rule from the Chebyshev polynomial equations and some empirical experiments. Following the principle presented in \cite{glorot2010understanding}, we searched for the initialisation scheme keeping the expected amplitude of the signal constant through the different layers of the neural network.
\end{mdframed}

\begin{mdframed}[style=comment]
Figure 10: Random selection of 24 learned filters from the fifth spherical convolutional layer L5. Top: section. Bottom: gnomonic projection.

- There is not enough information in this caption. Can you explain that there is some obvious peak structure. I think a colourbar on the gnomonic projection would be useful too.
\end{mdframed}
\todo{Assign: @all}
% \nati{The problem is that random filters are picky as well and look very similar. Should we say that? I do not think a colorbar is necessary since there are numbered axes on the top plots. Do you agree?}
\TK{Added sentence}

### 2019-03-04 scnn.tex
update comments some more

HEALPix is a particular case of a more general class of schemes based on a hierarchical subdivision of a base polyhedron.
Another example is the geodesic grids which are based on geodesic polyhedrons, i.e., polyhedrons made of triangular faces. A counter-example is the equirectangular projection, which is not constructed from a base polyhedron, although it can be subdivided.
In the particular HEALPix case, the base is a rhombic dodecahedron, i.e., a polyhedron made from 12 congruent rhombic faces.\footnote{A rhombus is a quadrilateral whose four sides all have the same length.}
See \figref{pooling} for an illustration of the base rhombic dodecahedron and its subdivisions.

@@

Fortunately, both of these computational issues are overcome by defining the convolution kernel $h$ as a polynomial $h_\theta(\lambda) = \sum_{k=0}^K \theta_k \lambda^k$ of degree $K$ parametrised by $K+1$ coefficients $\theta$.
The filtering operation \eqnref{graph_convolution_fourier} becomes
\begin{equation} \label{eqn:graph_convolution_monomial}
	h_\theta(\L) \f =  \U \left(\sum_{k=0}^K \theta_k \bLambda^k \right) \U\trans \f = \sum_{k=0}^K \theta_k \L^k \f,
\end{equation}
where $\L^k$ captures $k$-neighborhoods.\footnote{The 1-neighborhood of a vertex is the set of vertices that are directly connected to it. The $k$-neighborhood of a vertex is the set of vertices that are connected to it through paths of length $k$.}
The entry $(\L^k)_{ij}$ is the sum of all weighted paths of length $k$ between vertices $v_i$ and $v_j$, where the weight of a path is the multiplication of all the edge weights on the path.
Hence, it is non-zero if and only if vertices $v_i$ and $v_j$ are connected by at least one path of length $k$.
Filtering with a polynomial convolution kernel can thus be interpreted in the pixel (vertex) domain as a weighted linear combination of neighboring pixel values.
In the classical setting, convolutions are also weighted local sums.
The weights are however given by the filter coefficients only, and there is one coefficient per pixel in the patch.
In the graph setting, the weights are determined by the filter coefficients $\theta$ and the Laplacian $\L$, and there is one coefficient per neighborhood, not vertex.
That defines radial filters, values of which depend only on the distance to the center, and not on the direction.
% It is required as there is no way to order and match vertices from the same neighborhood.
% The number of vertices in a neighborhood most often even vary from vertex to vertex.
While it may seem odd to restrict filters to be 1D while the sphere is 2D, radial filters are direction-less and result in rotation equivariant convolutions.
We note that restricting the graph convolutional kernel to a polynomial is similar to restricting the classical Euclidean convolution to a fixed-size patch.
Similarly, each column of the matrix $\sum_{k=0}^K \theta_k \L^k$ defines an irregular patch of radius $K$.
Hence, filters designed as polynomials of the Laplacian have local support in the vertex domain.

@@

The layer $FC: \R^{F_{in}} \rightarrow \R^{F_{out}}$ is defined as
\begin{equation} \label{eqn:fc_layer}
	\y = FC(\x) = \b W \x + \b b ,
\end{equation}
where $\W \in \R^{F_{out} \times F_{in}}$ and $\b b \in \R^{F_{out}}$ are the parameters to be learned.
Note that the output $\Y \in \R^{N_{pix} \times F_{out}}$ of the last $GC$ (or the output $\Y \in \R^{N_{stat} \times F_{out}}$ of the $ST$), is vectorized as $\x = \vect(\X) \in \R^{F_{in}}$ before being fed to the first $FC$, where $F_{in} = N_{pix} \times F_{out}$ (or $N_{stat} \times F_{out}$).

The softmax layer is the last layer in a neural network engineered for classification. Given the output $\x \in \R^{N_{classes}}$ of the last $FC$ (or $ST$), called the logits in the deep learning literature, the softmax layer outputs $\y = SM(\x)$ such that
\begin{equation*}
	y_i = \frac{\exp(x_i)}{\sum_{j=1}^{N_{classes}} \exp(x_j)} \in [0, 1], \ \forall i \in [N_{classes}],
\end{equation*}
where $N_{classes}$ is the number of classes to discriminate.
Thanks to the softmax, the output $\y \in \R^{N_{classes}}$ of a neural network is a discretised conditional distribution for the class given the data and trained parameters.
That is, $y_i$ is the confidence of the network that the input sample belongs to class $i$.
% Note that in general, this probability distribution does not correspond to the true posterior distribution $p(y|x)$.
This last layer is actually normalizing $\x$ into $\y$ such that $\| \y \|_1 = \sum_i y_i = 1$.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.85\linewidth]{section_filter_last}
	\vspace{0.5cm}\\
	\includegraphics[width=0.85\linewidth]{gnonomic_filter_last}
	\caption{Random selection of 24 learned filters from the fifth spherical convolutional layer $L_5$. Top: section. Bottom: gnomonic projection.}
	\label{fig:learned_filter}
\end{figure}

>>>>

HEALPix is a particular case of a more general class of schemes based on a hierarchical subdivision of a base polyhedron.
Another example is the geodesic grids which are based on geodesic polyhedrons, i.e., polyhedrons made of triangular faces. A counter-example is the equirectangular projection, which is not constructed from a base polyhedron, although it can be subdivided.
In the particular HEALPix case, the base is a rhombic dodecahedron, i.e., a polyhedron made from 12 congruent rhombic faces.
% \footnote{A rhombus is a quadrilateral whose four sides all have the same length.}
See \figref{pooling} for an illustration of the base rhombic dodecahedron and its subdivisions.

@@

Fortunately, both of these computational issues are overcome by defining the convolution kernel $h$ as a polynomial $h_\theta(\lambda) = \sum_{k=0}^K \theta_k \lambda^k$ of degree $K$ parametrised by $K+1$ coefficients $\theta$.
The filtering operation \eqnref{graph_convolution_fourier} becomes
\begin{equation} \label{eqn:graph_convolution_monomial}
	h_\theta(\L) \f =  \U \left(\sum_{k=0}^K \theta_k \bLambda^k \right) \U\trans \f = \sum_{k=0}^K \theta_k \L^k \f,
\end{equation}
where $\L^k$ captures $k$-neighborhoods.\footnote{In the context of graph, one way to define the distance between two vertices is to count the number of edges in the shortest path between these two vertices. Hence, we define the 1-neighborhood of a vertex as the set of vertices that are directly connected to it. The $k$-neighborhood of a vertex is the set of vertices that are connected to it through paths of length $k$.}
The entry $(\L^k)_{ij}$ is the sum of all weighted paths of length $k$ between vertices $v_i$ and $v_j$, where the weight of a path is the multiplication of all the edge weights on the path.
Hence, it is non-zero if and only if vertices $v_i$ and $v_j$ are connected by at least one path of length $k$.
Filtering with a polynomial convolution kernel can thus be interpreted in the pixel (vertex) domain as a weighted linear combination of neighboring pixel values.
In the classical setting, convolutions are also weighted local sums.
The weights are however given by the filter coefficients only, and there is one coefficient per pixel in the patch.
In the graph setting, the weights are determined by the filter coefficients $\theta$ and the Laplacian $\L$, and there is one coefficient per neighborhood, not vertex.
That defines radial filters, values of which depend only on the distance to the center, and not on the direction.
% It is required as there is no way to order and match vertices from the same neighborhood.
% The number of vertices in a neighborhood most often even vary from vertex to vertex.
While it may seem odd to restrict filters to be 1D while the sphere is 2D, radial filters are direction-less and result in rotation equivariant convolutions.
We note that restricting the graph convolutional kernel to a polynomial is similar to restricting the classical Euclidean convolution to a fixed-size patch.
Similarly, each column of the matrix $\sum_{k=0}^K \theta_k \L^k$ defines an irregular patch of radius $K$.
Hence, filters designed as polynomials of the Laplacian have local support in the vertex domain.

@@

The layer $FC: \R^{F_{in}} \rightarrow \R^{F_{out}}$ is defined as
\begin{equation} \label{eqn:fc_layer}
	\y = FC(\x) = \b W \x + \b b ,
\end{equation}
where $\W \in \R^{F_{out} \times F_{in}}$ and $\b b \in \R^{F_{out}}$ are the parameters to be learned.
Note that the output $\Y \in \R^{N_{pix} \times F_{out}}$ of the last $GC$ (or the output $\Y \in \R^{N_{stat} \times F_{out}}$ of the statistical layer), is vectorized as $\x = \vect(\X) \in \R^{F_{in}}$ before being fed to the first $FC$, where $F_{in} = N_{pix} \times F_{out}$ (or $N_{stat} \times F_{out}$).

The softmax layer is the last layer in a neural network engineered for classification. Given the output $\x \in \R^{N_{classes}}$ of the last $FC$ (or statistical layer), called the logits in the deep learning literature, the softmax layer outputs $\y = SM(\x)$ such that
\begin{equation*}
	y_i = \frac{\exp(x_i)}{\sum_{j=1}^{N_{classes}} \exp(x_j)} \in [0, 1], \ \forall i \in [N_{classes}],
\end{equation*}
where $N_{classes}$ is the number of classes to discriminate.
Thanks to the softmax, the output $\y \in \R^{N_{classes}}$ of a neural network is a discretised conditional distribution for the class given the data and trained parameters.
That is, $y_i$ is the confidence of the network that the input sample belongs to class $i$.
% Note that in general, this probability distribution does not correspond to the true posterior distribution $p(y|x)$.
This last layer is actually normalizing $\x$ into $\y$ such that $\| \y \|_1 = \sum_i y_i = 1$.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.85\linewidth]{section_filter_last}
	\vspace{0.5cm}\\
	\includegraphics[width=0.85\linewidth]{gnonomic_filter_last}
	\caption{Random selection of 24 learned filters from the fifth spherical convolutional layer $L_5$. Top: section. Bottom: gnomonic projection.
	The structure in the filters resembles peaks, which is not unexpected, given that the convergence maps largely consist of concentrated clumps.}
	\label{fig:learned_filter}
\end{figure}
